### 2016-08-15 Documents/treatment_cascade.tex
Working on supplement.

\begin{longtable}{lrrrl}
  \hline
  & & & \multicolumn{1}{c}{Viral} \\
  Country & \multicolumn{1}{l}{Diagnosed}
  & \multicolumn{1}{l}{Treated} & \multicolumn{1}{l}{suppression}
  & References \\
  \hline
  Afghanistan & 5.0\% & 100.0\% & 84.0\% & \cite{Unaids2016-an} \\
  Algeria & 93.0\% & 91.7\% & 50.0\% & \cite{Unaids2016-an} \\
  Angola & 26.0\% & 100.0\% & 92.0\% & \cite{Unaids2016-an}, 2013a \\
  Argentina & 79.0\% & 79.3\% & 66.0\% & \cite{Unaids2016-an}, 2015a \\
  Armenia & 46.0\% & 55.4\% & 77.0\% & \cite{Unaids2016-an} \\
  Australia & 90.0\% & 99.4\% & 92.5\% & \cite{Unaids2016-an}, \url{https://kirby.unsw.edu.au/sites/default/files/hiv/resources/HIVASRsuppl2014_online.pdf} \\
  Austria & 80.0\% & 86.8\% & 63.0\% & 28th Report of the Austrian HIV Cohort Study (Innsbruck, September 30th, 2015), \url{https://www.google.com/url?sa=t&rct=j&q=&esrc=s&source=web&cd=12&ved=0ahUKEwjQkqHwwtLNAhXG0h4KHYxYDdM4ChAWCB0wAQ&url=http%3A%2F%2Fdermatologie.tirol-kliniken.at%2Fdata.cfm%3Fvpath%3Ddokumente%2F2015-hiv-kohortenbericht&usg=AFQjCNH9LyCKkigHECw7oYAEq_NBfRHjKw&sig2=eywFWu8eQN6aWf_QwHNLlw&bvm=bv.125801520,d.dmo&cad=rja} \\
  Azerbaijan & 35.0\% & 76.9\% & 67.0\% & \cite{Unaids2016-an} \\
  Bahamas & 37.0\% & 98.2\% & 69.0\% & \cite{Unaids2016-an} \\
  Bangladesh & 36.0\% & 41.6\% & 31.0\% & \cite{Unaids2016-an} \\
  Barbados & 85.0\% & 50.6\% & 88.0\% & \cite{Unaids2016-an} \\
  Belarus & 20.0\% & 100.0\% & 71.0\% & \cite{Unaids2016-an}, 2015b \\
  Belgium & 80.0\% & 74.5\% & 91.0\% & \cite{Unaids2016-an}, \url{http://ec.europa.eu/chafea/documents/health/hiv-cluster-meeting-material/presentations-pharis-s3_en.pdf}, 2015a \\
  Belize & 46.0\% & 99.0\% & 47.0\% & \cite{Unaids2016-an}, 2014b \\
  Benin & 48.0\% & 100.0\% & 93.0\% & \cite{Unaids2016-an}, 2011a \\
  Bhutan & 40.3\% & 41.4\% & 87.0\% & \cite{Unaids2016-an, Unaids2016-am}, 2014b \\
  Bolivia & 29.0\% & 97.7\% & 69.0\% & \cite{Unaids2016-an} \\
  Botswana & 80.0\% & 97.2\% & 90.0\% & \cite{Unaids2016-an} \\
  Brazil & 55.0\% & 100.0\% & 77.0\% & \cite{Unaids2016-an}, 2014b \\
  Bulgaria & 21.8\% & 72.7\% & 81.6\% & \cite{Unaids2016-an, Unaids2016-am}\\
  Burkina Faso & 83.0\% & 66.3\% & 79.0\% & \cite{Unaids2016-an}, 2013? \\
  Burundi & 54.0\% & 100.0\% & 90.0\% & \cite{Unaids2016-an}, 2015b \\
  Cambodia & 82.0\% & 89.0\% & 94.0\% & \cite{Unaids2016-an} \\
  Cameroon & 27.0\% & 100.0\% & 60.0\% & \cite{Unaids2016-an}, 2015a \\
  Canada & 79.0\% & 64.6\% & 68.6\% & \url{http://www.catie.ca/sites/default/files/2014-HIV-Estimates-in-Canada-EN.pdf}, \url{http://www.catie.ca/en/catienews/2014-02-19/gaps-british-columbia-s-hiv-treatment-cascade} \\
  Cape Verde & 29.0\% & 100.0\% & 95.0\% & \cite{Unaids2016-an}, 2015b \\
  Central African Republic & 24.0\% & 98.1\% & 91.0\% & \cite{Unaids2016-an}, 2013? \\
  Chad & 36.0\% & 100.0\% & 91.0\% & \cite{Unaids2016-an}, 2013? \\
  Chile & 90.0\% & 96.7\% & 95.0\% & \cite{Unaids2016-an}, 2015b \\
  China & 64.2\% & 59.0\% & 91.0\% & \cite{Unaids2016-an, AIDSdatahub-fg}, \url{http://www.avert.org/professionals/hiv-around-world/asia-pacific/china} \\
  Colombia & 45.0\% & 73.0\% & 70.0\% & \url{http://ecdc.europa.eu/en/publications/Publications/dublin-declaration-continuum-of-care-2014.pdf} \\
  Republic of Congo & 37.4\% & 50.5\% & 79.4\% & \cite{Unaids2016-am} \\
  Costa Rica & 56.0\% & 100.0\% & 60.7\% & \cite{Unaids2016-an} \\
  Cuba & 64.0\% & 100.0\% & 59.0\% & \cite{Unaids2016-an} \\
  Côte d'Ivoire & 51.0\% & 68.7\% & 69.0\% & \cite{Unaids2016-an}, 2014b \\
  DR Congo & 33.0\% & 100.0\% & 86.0\% & \cite{Unaids2016-an}, 2015a \\
  Denmark & 85.0\% & 72.9\% & 95.2\% & \url{http://www.aidsmap.com/Australia-performs-best-in-HIV-treatment-cascade-62-with-undetectable-viral-load/page/2919074/} \\
  Djibouti & 70.0\% & 31.4\% & 86.0\% & \cite{Unaids2016-an}, 2015b \\
  Dominican Republic & 47.0\% & 98.4\% & 95.0\% & \cite{Unaids2016-an}, 2015b \\
  Ecuador & 53.0\% & 97.4\% & 85.0\% & \cite{Unaids2016-an}, 2012b \\
  Egypt & 40.6\% & 56.5\% & 80.0\% & \url{http://english.alarabiya.net/en/perspective/analysis/2014/08/28/Registered-HIV-AIDS-cases-rising-in-Egypt-U-N-.html}, 2015b \\
  El Salvador & 54.0\% & 100.0\% & 83.0\% & \cite{Unaids2016-an}, 2015b \\
  Equatorial Guinea & 31.0\% & 100.0\% & 82.0\% & \cite{Unaids2016-an} \\
  Eritrea & 60.0\% & 100.0\% & 77.0\% & \cite{Unaids2016-an}, 2015b \\
  Estonia & 87.0\% & 33.3\% & 65.5\% & \url{https://www.sm.ee/sites/default/files/content-editors/eesmargid_ja_tegevused/Tervis/Tervislik_eluviis/who_raport_hiv_aids_treatment_and_care_in_estonia.pdf} \\
  Ethiopia & 71.5\% & 70.0\% & 82.0\% & \cite{Unaids2016-am}, \url{https://dhsprogram.com/pubs/pdf/CR30/CR30.pdf}, 2012b \\
  France & 81.0\% & 91.4\% & 70.3\% & \url{http://www.aidsmap.com/Australia-performs-best-in-HIV-treatment-cascade-62-with-undetectable-viral-load/page/2919074/} \\
  Gabon & 85.0\% & 68.6\% & 64.0\% & \cite{Unaids2016-an}, 2014b \\
  Gambia & 24.0\% & 100.0\% & 83.0\% & \cite{Unaids2016-an}, 2015a \\
  Georgia & 43.0\% & 72.5\% & 84.0\% & \cite{Unaids2016-an} \\
  Ghana & 34.0\% & 100.0\% & 92.0\% & \cite{Unaids2016-an}, 2014b \\
  Greece & 90.2\% & 52.8\% & 95.0\% & \cite{Unaids2016-an, Unaids2016-am}, 2015b \\
  Guatemala & 38.0\% & 76.8\% & 83.0\% & \cite{Unaids2016-an} \\
  Guinea & 32.0\% & 90.9\% & 65.0\% & \cite{Unaids2016-an} \\
  Guyana & 58.0\% & 99.7\% & 80.0\% & \cite{Unaids2016-an} \\
  Haiti & 54.0\% & 96.3\% & 76.0\% & \cite{Unaids2016-an}, 2013b \\
  Honduras & 56.0\% & 92.9\% & 83.0\% & \cite{Unaids2016-an}, 2015b \\
  India & 95.2\% & 41.5\% & 73.0\% & \cite{AIDSdatahub-fg}, 2014b \\
  Indonesia & 9.0\% & 98.6\% & 71.0\% & \cite{Unaids2016-an}, 2014a \\
  Iran & 31.0\% & 28.4\% & 67.0\% & \cite{Unaids2016-an} \\
  Italy & 70.0\% & 87.6\% & 62.4\% & \url{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4348082/} \\
  Jamaica & 33.0\% & 95.1\% & 67.0\% & \cite{Unaids2016-an} \\
  Kazakhstan & 72.0\% & 35.7\% & 60.0\% & \cite{Unaids2016-an} \\
  Kenya & 58.0\% & 100.0\% & 84.0\% & \cite{Unaids2016-an} \\
  Kyrgyzstan & 50.0\% & 43.0\% & 62.0\% & \cite{Unaids2016-an} \\
  Laos & 62.0\% & 45.8\% & 83.0\% & \cite{AIDSdatahub-fg}, 2015b \\
  Lebanon & 36.0\% & 100.0\% & 79.0\% & \cite{Unaids2016-an} \\
  Lesotho & 67.0\% & 62.6\% & 68.0\% & \cite{Unaids2016-an} \\
  Liberia & 41.0\% & 65.7\% & 74.0\% & \cite{Unaids2016-an}, 2013b \\
  Madagascar & 4.0\% & 69.2\% & 73.0\% & \cite{Unaids2016-an}, 2012b \\
  Malawi & 61.0\% & 100.0\% & 78.0\% & \cite{Unaids2016-an}, 2012b \\
  Malaysia & 88.1\% & 24.6\% & 85.0\% & \cite{Unaids2016-an, AIDSdatahub-fg} \\
  Mali & 71.0\% & 41.4\% & 72.0\% & \cite{Unaids2016-an}, 2014a \\
  Mauritania & 18.0\% & 100.0\% & 87.0\% & \cite{Unaids2016-an} \\
  Mauritius & 65.0\% & 47.7\% & 82.0\% & \cite{Unaids2016-an}, 2014b \\
  Mexico & 58.0\% & 99.7\% & 85.0\% & \cite{Unaids2016-an}, 2015b \\
  Moldova & 41.0\% & 50.7\% & 78.0\% & \cite{Unaids2016-an} \\
  Mongolia & 41.0\% & 39.8\% & 93.0\% & \cite{Unaids2016-an, Unaids2016-am} \\
  Morocco & 46.0\% & 76.4\% & 82.0\% & \cite{Unaids2016-an}, 2015a \\
  Mozambique & 53.0\% & 99.5\% & 66.0\% & \cite{Unaids2016-an}, 2015b \\
  Myanmar & 47.0\% & 96.1\% & 87.0\% & \cite{Unaids2016-an} \\
  Namibia & 69.0\% & 97.5\% & 85.0\% & \cite{Unaids2016-an}, 2015b \\
  Nepal & 57.0\% & 50.9\% & 91.0\% & \cite{Unaids2016-an} \\
  Netherlands & 73.0\% & 93.2\% & 89.8\% & \url{http://www.aidsmap.com/Australia-performs-best-in-HIV-treatment-cascade-62-with-undetectable-viral-load/page/2919074/} \\
  New Zealand & 85.0\% & 80.0\% & 98.4\% & \cite{Unaids2016-am} \\
  Nicaragua & 47.0\% & 70.2\% & 61.0\% & \cite{Unaids2016-an} \\
  Niger & 28.0\% & 99.4\% & 67.0\% & \cite{Unaids2016-an}, 2015b \\
  Nigeria & 58.0\% & 41.4\% & 81.0\% & \cite{Unaids2016-an}, 2013b \\
  Pakistan & 6.0\% & 100.0\% & 58.0\% & \cite{Unaids2016-an} \\
  Panama & 59.0\% & 100.0\% & 66.0\% & \cite{Unaids2016-an} \\
  Papua New Guinea & 81.0\% & 64.2\% & 90.0\% & \cite{Unaids2016-an}, 2015a \\
  Paraguay & 71.0\% & 44.1\% & 70.0\% & \cite{Unaids2016-an} \\
  Peru & 64.0\% & 87.3\% & 82.0\% & \cite{Unaids2016-an}, 2015b \\
  Philippines & 63.0\% & 42.9\% & 92.0\% & \cite{Unaids2016-an} \\
  Russia & 49.0\% & 23.5\% & 80.0\% & \url{http://www.jiasociety.org/index.php/jias/article/view/19506} \\
  Rwanda & 85.0\% & 92.9\% & 94.0\% & \cite{Unaids2016-an}, 2012b \\
  Saint Lucia & 49.7\% & 70.2\% & 71.0\% & \cite{Unaids2016-an, Unaids2016-am} \\
  Senegal & 56.0\% & 73.2\% & 77.0\% & \cite{Unaids2016-an}, 2015b \\
  Sierra Leone & 27.0\% & 100.0\% & 71.0\% & \cite{Unaids2016-an}, 2014a \\
  Somalia & 8.0\% & 100.0\% & 81.0\% & \cite{Unaids2016-an}, 2015a \\
  South Africa & 57.0\% & 84.0\% & 71.0\% & \cite{Unaids2016-an}, 2013b \\
  South Sudan & 11.0\% & 100.0\% & 76.0\% & \cite{Unaids2016-an}, 2015b \\
  Spain & 74.0\% & 98.7\% & 89.0\% & \cite{Unaids2016-an} \\
  Sri Lanka & 45.0\% & 41.6\% & 81.0\% & \cite{Unaids2016-an}, 2015b \\
  Sudan & 8.0\% & 96.0\% & 90.0\% & \cite{Unaids2016-an}, 2012b \\
  Suriname & 54.0\% & 100.0\% & 76.0\% & \cite{Unaids2016-an}, 2014b \\
  Swaziland & 67.0\% & 100.0\% & 51.0\% & \cite{Unaids2016-an} \\
  Sweden & 90.0\% & 92.1\% & 94.7\% & \url{http://www.croiconference.org/sessions/swedish-hiv-treatement-cascade} \\
  Switzerland & 80.9\% & 87.8\% & 96.3\% & \url{http://www.natap.org/2015/IAS/Jake_IAS_Slides_July20X.pdf} \\
  Tajikistan & 34.0\% & 48.5\% & 77.0\% & \cite{Unaids2016-an} \\
  Tanzania & 53.0\% & 99.9\% & 67.0\% & \cite{Unaids2016-an}, 2015b \\
  Thailand & 83.0\% & 78.3\% & 96.0\% & \cite{Unaids2016-an} \\
  Timor-Leste & 43.8\% & 76.8\% & 82.0\% & \cite{Unaids2016-am} \\
  Togo & 41.0\% & 100.0\% & 87.0\% & \cite{Unaids2016-an}, 2014b \\
  Trinidad and Tobago & 62.0\% & 96.6\% & 90.0\% & \cite{Unaids2016-an}, 2015b \\
  Tunisia & 61.0\% & 44.3\% & 81.0\% & \cite{Unaids2016-an}, 2014b \\
  Uganda & 63.0\% & 87.9\% & 77.0\% & \cite{Unaids2016-an}, 2015b \\
  Ukraine & 27.0\% & 98.2\% & 86.0\% & \cite{Unaids2016-an}, 2015a \\
  United Kingdom & 82.5\% & 89.8\% & 94.7\% & \url{https://www.gov.uk/government/uploads/system/uploads/attachment_data/file/477702/HIV_in_the_UK_2015_report.pdf} \\
  United States & 87.0\% & 52.0\% & 81.9\% & \url{http://www.cdc.gov/hiv/pdf/surveillance_Report_vol_19_no_3.pdf}, \url{http://www.cdc.gov/hiv/statistics/overview/geographicdistribution.html} \\
  Uruguay & 60.0\% & 100.0\% & 93.0\% & \cite{Unaids2016-an}, 2014b \\
  Uzbekistan & 60.0\% & 44.8\% & 88.0\% & \cite{Unaids2016-an}, 2015b \\
  Venezuela & 58.0\% & 97.5\% & 95.0\% & \cite{Unaids2016-an}, 2015a \\
  Vietnam & 80.0\% & 50.8\% & 71.0\% & \cite{Unaids2016-an}, 2014b \\
  Yemen & 14.0\% & 98.6\% & 77.0\% & \cite{Unaids2016-an}, 2013a \\
  Zambia & 73.0\% & 86.3\% & 93.0\% & \cite{Unaids2016-an}, 2015a \\
  Zimbabwe & 64.0\% & 96.9\% & 84.0\% & \cite{Unaids2016-an} \\
  \hline
\end{longtable}
2011a: Viral suppression is 12-month retention on treatment, all ages, for 2011.
2012b: Viral suppression is 12-month retention on treatment, ages 15+, for 2012.
2013a: Viral suppression is 12-month retention on treatment, all ages, for 2013.
2013b: Viral suppression is 12-month retention on treatment, ages 15+, for 2013.
2014a: Viral suppression is 12-month retention on treatment, all ages, for 2014.
2014b: Viral suppression is 12-month retention on treatment, ages 15+, for 2014.
2015a: Viral suppression is 12-month retention on treatment, all ages, for 2015.
2015b: Viral suppression is 12-month retention on treatment, ages 15+, for 2015.