### 2016-08-15 Documents/prevalence.tex
Working on supplement tables.

\begin{longtable}{lrrrrrrrrrrrrrrrrrrrrrrrrrrl}
  \caption{prevalence.}
  \label{data_source_prevalence}
  \\[-2ex]
  \hline
  Country & 1990 & 1991 & 1992 & 1993 & 1994 & 1995 & 1996 & 1997 & 1998 & 1999 & 2000 & 2001 & 2002 & 2003 & 2004 & 2005 & 2006 & 2007 & 2008 & 2009 & 2010 & 2011 & 2012 & 2013 & 2014 & 2015 & References\\
  \hline
  \endfirsthead
  \caption{(continued)}
  \\[-2ex]
  \hline
  Country & 1990 & 1991 & 1992 & 1993 & 1994 & 1995 & 1996 & 1997 & 1998 & 1999 & 2000 & 2001 & 2002 & 2003 & 2004 & 2005 & 2006 & 2007 & 2008 & 2009 & 2010 & 2011 & 2012 & 2013 & 2014 & 2015 & References\\
  \hline
  \endhead
  \hline
  \endfoot
  \endlastfoot
  Afghanistan & 0.0077\% & 0.0095\% & 0.01\% & 0.012\% & 0.013\% & 0.014\% & 0.014\% & 0.015\% & 0.016\% & 0.017\% & 0.018\% & 0.019\% & 0.02\% & 0.02\% & 0.022\% & 0.023\% & 0.025\% & 0.027\% & 0.028\% & 0.03\% & 0.032\% & 0.033\% & 0.034\% & 0.035\% & 0.036\% & --- & \url{http://www.unaids.org/sites/default/files/country/documents/AFG_narrative_report_2014.pdf }, Estimated PLHIV (total) / Population (total)\\
  Algeria & --- & --- & --- & --- & 0.0079\% & 0.0083\% & 0.0094\% & 0.01\% & 0.011\% & 0.013\% & 0.014\% & 0.015\% & 0.017\% & 0.018\% & 0.02\% & 0.021\% & 0.023\% & 0.026\% & 0.028\% & 0.03\% & 0.033\% & 0.035\% & 0.037\% & 0.038\% & 0.039\% & 0.04\% & \url{http://aidsinfo.unaids.org/}, PLHIV 15+ / population 15-49\\
  Angola & 0.5\% & 0.5\% & 0.6\% & 0.7\% & 0.8\% & 0.9\% & 1.1\% & 1.2\% & 1.3\% & 1.4\% & 1.5\% & 1.6\% & 1.7\% & 1.8\% & 1.8\% & 1.9\% & 2.0\% & 2.0\% & 2.1\% & 2.1\% & 2.1\% & 2.2\% & 2.2\% & 2.2\% & 2.2\% & 2.2\% & \url{http://aidsinfo.unaids.org/}\\
  Argentina & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & \url{http://aidsinfo.unaids.org/}\\
  Armenia & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & \url{http://aidsinfo.unaids.org/}\\
  Australia & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & \url{http://aidsinfo.unaids.org/}\\
  Austria & --- & --- & --- & --- & --- & --- & --- & 0.062\% & 0.076\% & 0.093\% & 0.11\% & 0.13\% & 0.16\% & 0.19\% & 0.22\% & 0.24\% & 0.28\% & 0.31\% & 0.35\% & 0.39\% & 0.4\% & 0.44\% & 0.47\% & --- & --- & --- & \url{http://www.unaids.org/en/resources/campaigns/globalreport2013/globalreport}\\
  Azerbaijan & --- & --- & --- & --- & --- & --- & --- & --- & 0.025\% & 0.032\% & 0.04\% & 0.048\% & 0.055\% & 0.062\% & 0.071\% & 0.08\% & 0.09\% & 0.1\% & 0.11\% & 0.13\% & 0.14\% & 0.15\% & 0.17\% & 0.18\% & 0.21\% & --- & \url{http://aidsinfo.unaids.org/}, PLHIV 15+ / population 15-49\\
  Bahamas & 1.6\% & 1.8\% & 2.0\% & 2.3\% & 2.5\% & 2.7\% & 2.8\% & 2.9\% & 3.0\% & 3.0\% & 3.1\% & 3.1\% & 3.1\% & 3.1\% & 3.1\% & 3.1\% & 3.1\% & 3.1\% & 3.2\% & 3.2\% & 3.2\% & 3.2\% & 3.2\% & 3.2\% & 3.2\% & 3.2\% & \url{http://aidsinfo.unaids.org/}\\
  Bangladesh & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & 0.0016\% & 0.0022\% & 0.003\% & 0.0038\% & 0.0048\% & 0.0058\% & 0.0065\% & 0.0074\% & 0.0083\% & 0.009\% & 0.0097\% & 0.01\% & 0.01\% & 0.011\% & 0.011\% & 0.011\% & \url{http://aidsinfo.unaids.org/}, Estimated PLHIV (adult) / Population (15-49)\\
  Barbados & 0.2\% & 0.2\% & 0.2\% & 0.3\% & 0.3\% & 0.4\% & 0.4\% & 0.5\% & 0.5\% & 0.6\% & 0.7\% & 0.7\% & 0.8\% & 0.8\% & 0.9\% & 1\% & 1\% & 1.1\% & 1.2\% & 1.3\% & 1.3\% & 1.4\% & 1.4\% & 1.5\% & 1.5\% & 1.6\% & \url{http://aidsinfo.unaids.org/}\\
  Belarus & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & 0.1\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.4\% & 0.4\% & 0.4\% & 0.5\% & 0.6\% & 0.6\% & \url{http://aidsinfo.unaids.org/}\\
  Belgium & --- & --- & 0.045\% & 0.057\% & 0.072\% & 0.084\% & 0.097\% & 0.11\% & 0.12\% & 0.14\% & 0.16\% & 0.18\% & 0.19\% & 0.21\% & 0.23\% & 0.25\% & 0.28\% & 0.3\% & 0.33\% & 0.34\% & 0.36\% & 0.38\% & 0.41\% & --- & --- & --- & \url{http://www.unaids.org/en/resources/campaigns/globalreport2013/globalreport}, Estimated PLHIV (total) / Population (15-49)\\
  Belize & $<$0.1\% & $<$0.1\% & $<$0.1\% & 0.1\% & 0.2\% & 0.2\% & 0.3\% & 0.5\% & 0.6\% & 0.9\% & 1.1\% & 1.4\% & 1.7\% & 1.9\% & 2.0\% & 2.0\% & 1.9\% & 1.9\% & 1.8\% & 1.7\% & 1.7\% & 1.7\% & 1.6\% & 1.6\% & 1.6\% & 1.5\% & \url{http://aidsinfo.unaids.org/}\\
  Benin & 0.2\% & 0.3\% & 0.4\% & 0.6\% & 0.8\% & 0.9\% & 1.1\% & 1.2\% & 1.3\% & 1.4\% & 1.5\% & 1.5\% & 1.5\% & 1.5\% & 1.4\% & 1.4\% & 1.3\% & 1.3\% & 1.2\% & 1.2\% & 1.2\% & 1.2\% & 1.1\% & 1.1\% & 1.1\% & 1.1\% & \url{http://aidsinfo.unaids.org/}\\
  Bhutan & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.13\% & --- & --- & \url{https://www.cia.gov/library/publications/the-world-factbook/rankorder/2155rank.html}, \url{http://www.unaids.org/en/resources/campaigns/globalreport2013/globalreport}\\
  Bolivia & 0.2\% & 0.2\% & 0.2\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & \url{http://aidsinfo.unaids.org/}\\
  Botswana & 6.0\% & 8.4\% & 11.2\% & 14.2\% & 17.2\% & 19.9\% & 22.2\% & 24.0\% & 25.1\% & 25.8\% & 26.0\% & 26.0\% & 25.7\% & 25.3\% & 24.9\% & 24.6\% & 24.3\% & 24.1\% & 23.9\% & 23.6\% & 23.5\% & 23.3\% & 23.1\% & 22.8\% & 22.5\% & 22.2\% & \url{http://aidsinfo.unaids.org/}\\
  Brazil & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.6\% & 0.6\% & 0.6\% & 0.6\% & \url{http://aidsinfo.unaids.org/}\\
  Bulgaria & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & 0.042\% & 0.051\% & 0.058\% & 0.065\% & 0.073\% & 0.081\% & 0.09\% & 0.097\% & 0.1\% & 0.11\% & 0.12\% & 0.12\% & --- & --- & --- & \url{http://www.unaids.org/en/resources/campaigns/globalreport2013/globalreport}\\
  Burkina Faso & 3.4\% & 3.6\% & 3.7\% & 3.8\% & 3.7\% & 3.6\% & 3.4\% & 3.2\% & 2.9\% & 2.7\% & 2.4\% & 2.2\% & 1.9\% & 1.7\% & 1.6\% & 1.4\% & 1.3\% & 1.2\% & 1.1\% & 1.1\% & 1\% & 1\% & 0.9\% & 0.9\% & 0.9\% & 0.8\% & \url{http://aidsinfo.unaids.org/}\\
  Burundi & 0.9\% & 1.1\% & 1.4\% & 1.8\% & 2.1\% & 2.5\% & 2.7\% & 2.9\% & 3.1\% & 3.2\% & 3.2\% & 3.1\% & 2.9\% & 2.7\% & 2.5\% & 2.3\% & 2.2\% & 2.0\% & 1.8\% & 1.6\% & 1.5\% & 1.4\% & 1.3\% & 1.2\% & 1.1\% & 1\% & \url{http://aidsinfo.unaids.org/}\\
  Cambodia & --- & 0.1\% & 0.3\% & 0.6\% & 1\% & 1.4\% & 1.7\% & 1.8\% & 1.9\% & 1.8\% & 1.8\% & 1.6\% & 1.5\% & 1.4\% & 1.2\% & 1.1\% & 1.1\% & 1\% & 0.9\% & 0.9\% & 0.8\% & 0.8\% & 0.7\% & 0.7\% & 0.7\% & 0.6\% & \url{http://aidsinfo.unaids.org/}\\
  Cameroon & 2.0\% & 2.5\% & 3.0\% & 3.5\% & 3.9\% & 4.3\% & 4.7\% & 5.0\% & 5.2\% & 5.3\% & 5.4\% & 5.4\% & 5.4\% & 5.3\% & 5.3\% & 5.2\% & 5.1\% & 5.0\% & 4.9\% & 4.9\% & 4.8\% & 4.8\% & 4.7\% & 4.6\% & 4.6\% & 4.5\% & \url{http://aidsinfo.unaids.org/}\\
  Canada & 0.22\% & 0.22\% & 0.24\% & 0.24\% & 0.25\% & 0.25\% & 0.25\% & 0.26\% & 0.27\% & 0.27\% & 0.28\% & 0.29\% & 0.3\% & 0.32\% & 0.33\% & 0.34\% & 0.35\% & 0.36\% & 0.38\% & 0.39\% & 0.4\% & 0.43\% & --- & --- & 0.45\% & --- & \url{http://www.phac-aspc.gc.ca/aids-sida/publication/epi/2010/1-eng.php#a0502}, Estimated \# HIV infections (total) / Population (15-49)\\
  Cape Verde & 0.8\% & 0.9\% & 1\% & 1.1\% & 1.2\% & 1.3\% & 1.4\% & 1.5\% & 1.5\% & 1.6\% & 1.6\% & 1.6\% & 1.5\% & 1.5\% & 1.4\% & 1.3\% & 1.3\% & 1.2\% & 1.2\% & 1.1\% & 1.1\% & 1\% & 1\% & 1\% & 1\% & 1\% & \url{http://aidsinfo.unaids.org/}\\
  Central African Republic & 4.0\% & 4.8\% & 5.6\% & 6.4\% & 7.2\% & 7.8\% & 8.3\% & 8.7\% & 8.8\% & 8.8\% & 8.7\% & 8.4\% & 8.0\% & 7.5\% & 7.0\% & 6.5\% & 6.1\% & 5.7\% & 5.3\% & 5.0\% & 4.8\% & 4.5\% & 4.3\% & 4.1\% & 3.9\% & 3.7\% & \url{http://aidsinfo.unaids.org/}\\
  Chad & 1.2\% & 1.3\% & 1.5\% & 1.7\% & 1.9\% & 2.1\% & 2.4\% & 2.6\% & 2.8\% & 3.0\% & 3.2\% & 3.3\% & 3.4\% & 3.4\% & 3.4\% & 3.3\% & 3.2\% & 3.1\% & 3.0\% & 2.8\% & 2.7\% & 2.6\% & 2.4\% & 2.3\% & 2.2\% & 2.0\% & \url{http://aidsinfo.unaids.org/}\\
  Chile & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.3\% & 0.3\% & 0.3\% & \url{http://aidsinfo.unaids.org/}\\
  China & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & 0.086\% & --- & 0.091\% & --- & 0.095\% & --- & 0.1\% & --- & 0.11\% & 0.066\% & --- & \url{http://www.unaids.org/sites/default/files/country/documents/CHN_narrative_report_2015.pdf}, \url{http://www.ncbi.nlm.nih.gov/pmc/articles/PMC4730421/pdf/ijerph-13-00030.pdf}\\
  Colombia & 0.1\% & 0.1\% & 0.1\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.5\% & 0.5\% & \url{http://aidsinfo.unaids.org/}\\
  Congo & 3.9\% & 4.3\% & 4.8\% & 5.1\% & 5.4\% & 5.5\% & 5.6\% & 5.7\% & 5.6\% & 5.5\% & 5.4\% & 5.2\% & 4.9\% & 4.7\% & 4.4\% & 4.1\% & 3.9\% & 3.7\% & 3.5\% & 3.3\% & 3.2\% & 3.1\% & 3.0\% & 2.9\% & 2.8\% & --- & \url{http://www.who.int/gho/database/en/}\\
  Costa Rica & $<$0.1\% & $<$0.1\% & $<$0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & \url{http://aidsinfo.unaids.org/}\\
  Cuba & --- & --- & --- & --- & 0.016\% & 0.02\% & 0.023\% & 0.026\% & 0.03\% & 0.034\% & 0.039\% & 0.046\% & 0.054\% & 0.062\% & 0.073\% & 0.085\% & 0.098\% & 0.11\% & 0.13\% & 0.15\% & 0.18\% & 0.2\% & 0.24\% & 0.29\% & 0.33\% & 0.39\% & \url{http://aidsinfo.unaids.org/}, PLHIV 15+/ Population 15-49\\
  Côte d'Ivoire & 3.0\% & 3.5\% & 3.9\% & 4.3\% & 4.7\% & 5.1\% & 5.4\% & 5.6\% & 5.7\% & 5.7\% & 5.6\% & 5.5\% & 5.3\% & 5.1\% & 4.9\% & 4.6\% & 4.4\% & 4.2\% & 4.1\% & 3.9\% & 3.8\% & 3.7\% & 3.6\% & 3.4\% & 3.3\% & 3.2\% & \url{http://aidsinfo.unaids.org/}\\
  DR Congo & 1.7\% & 1.7\% & 1.7\% & 1.7\% & 1.7\% & 1.8\% & 1.8\% & 1.9\% & 1.9\% & 1.9\% & 1.9\% & 1.9\% & 1.9\% & 1.8\% & 1.8\% & 1.7\% & 1.6\% & 1.5\% & 1.4\% & 1.3\% & 1.2\% & 1.1\% & 1\% & 1\% & 0.9\% & 0.8\% & \url{http://aidsinfo.unaids.org/}\\
  Denmark & 0.049\% & 0.056\% & 0.06\% & 0.066\% & 0.071\% & 0.077\% & 0.085\% & 0.09\% & 0.1\% & 0.11\% & 0.13\% & 0.14\% & 0.15\% & 0.16\% & 0.17\% & 0.18\% & 0.19\% & 0.2\% & 0.21\% & 0.22\% & 0.23\% & 0.24\% & 0.26\% & --- & --- & --- & \url{http://www.unaids.org/en/resources/campaigns/globalreport2013/globalreport}\\
  Djibouti & 0.2\% & 0.3\% & 0.5\% & 0.8\% & 1.1\% & 1.5\% & 1.8\% & 2.2\% & 2.4\% & 2.7\% & 2.8\% & 2.8\% & 2.8\% & 2.7\% & 2.6\% & 2.5\% & 2.3\% & 2.2\% & 2.1\% & 1.9\% & 1.9\% & 1.8\% & 1.7\% & 1.6\% & 1.6\% & 1.6\% & \url{http://aidsinfo.unaids.org/}\\
  Dominican Republic & 0.6\% & 0.8\% & 1\% & 1.2\% & 1.5\% & 1.7\% & 1.9\% & 2.1\% & 2.2\% & 2.3\% & 2.3\% & 2.3\% & 2.2\% & 2.1\% & 2.0\% & 1.9\% & 1.8\% & 1.7\% & 1.5\% & 1.4\% & 1.4\% & 1.3\% & 1.2\% & 1.1\% & 1.1\% & 1\% & \url{http://aidsinfo.unaids.org/}\\
  Ecuador & 0.2\% & 0.2\% & 0.2\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & \url{http://aidsinfo.unaids.org/}\\
  Egypt & --- & --- & --- & --- & --- & --- & --- & 0.0035\% & 0.0037\% & 0.0043\% & 0.005\% & 0.0054\% & 0.0058\% & 0.0065\% & 0.007\% & 0.0079\% & 0.0085\% & 0.0093\% & 0.01\% & 0.011\% & 0.012\% & 0.013\% & 0.014\% & --- & 0.02\% & --- & \url{https://www.cia.gov/library/publications/the-world-factbook/rankorder/2155rank.html}, Estimated PLHIV (total) / Population (15-49)\\
  El Salvador & 0.1\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.3\% & 0.3\% & 0.3\% & 0.4\% & 0.4\% & 0.4\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.6\% & 0.6\% & 0.6\% & 0.6\% & 0.6\% & 0.6\% & 0.6\% & 0.5\% & 0.5\% & 0.5\% & \url{http://aidsinfo.unaids.org/}\\
  Equatorial Guinea & 0.7\% & 0.7\% & 0.8\% & 0.9\% & 1\% & 1.1\% & 1.3\% & 1.5\% & 1.8\% & 2.1\% & 2.5\% & 3.0\% & 3.5\% & 4.1\% & 4.6\% & 5.2\% & 5.7\% & 6.1\% & 6.4\% & 6.5\% & 6.5\% & 6.3\% & 6.0\% & 5.6\% & 5.3\% & 4.9\% & \url{http://aidsinfo.unaids.org/}\\
  Eritrea & 0.5\% & 0.6\% & 0.8\% & 1\% & 1.2\% & 1.3\% & 1.5\% & 1.5\% & 1.6\% & 1.6\% & 1.6\% & 1.5\% & 1.4\% & 1.3\% & 1.2\% & 1.1\% & 1\% & 0.9\% & 0.9\% & 0.8\% & 0.8\% & 0.7\% & 0.7\% & 0.7\% & 0.6\% & 0.6\% & \url{http://aidsinfo.unaids.org/}\\
  Estonia & --- & --- & --- & --- & --- & --- & --- & --- & 0.56\% & 0.73\% & 0.98\% & 1.3\% & 1.6\% & 1.8\% & 2.0\% & 2.2\% & 2.4\% & 2.5\% & 2.6\% & 2.6\% & 2.8\% & 2.9\% & 2.9\% & --- & --- & --- & \url{http://www.unaids.org/en/resources/campaigns/globalreport2013/globalreport}\\
  Ethiopia & 1.3\% & 1.7\% & 2.1\% & 2.5\% & 2.8\% & 3.1\% & 3.3\% & 3.4\% & 3.4\% & 3.4\% & 3.2\% & 3.1\% & 2.8\% & 2.6\% & 2.4\% & 2.1\% & 1.9\% & 1.7\% & 1.6\% & 1.5\% & 1.4\% & 1.3\% & 1.3\% & 1.2\% & 1.2\% & --- & \url{https://www.cia.gov/library/publications/the-world-factbook/rankorder/2155rank.html}, \url{http://data.worldbank.org/indicator/SH.DYN.AIDS.ZS}\\
  France & 0.31\% & 0.31\% & 0.3\% & 0.29\% & 0.29\% & 0.29\% & 0.29\% & 0.32\% & 0.33\% & 0.36\% & 0.36\% & 0.38\% & 0.39\% & 0.39\% & 0.42\% & 0.42\% & 0.44\% & 0.44\% & 0.46\% & 0.48\% & 0.49\% & 0.51\% & --- & --- & 0.51\% & --- & \url{http://www.aidsmap.com/Australia-performs-best-in-HIV-treatment-cascade-62-with-undetectable-viral-load/page/2919074/}, \url{http://www.unaids.org/en/resources/campaigns/globalreport2013/globalreport}\\
  Gabon & 0.9\% & 1.2\% & 1.5\% & 1.9\% & 2.3\% & 2.8\% & 3.3\% & 3.7\% & 4.2\% & 4.6\% & 5.0\% & 5.2\% & 5.4\% & 5.4\% & 5.5\% & 5.4\% & 5.4\% & 5.2\% & 5.0\% & 4.8\% & 4.6\% & 4.4\% & 4.2\% & 4.0\% & 3.9\% & 3.8\% & \url{http://aidsinfo.unaids.org/}\\
  Gambia & 0.2\% & 0.2\% & 0.3\% & 0.4\% & 0.5\% & 0.6\% & 0.8\% & 1\% & 1.3\% & 1.5\% & 1.7\% & 1.9\% & 2.0\% & 2.1\% & 2.2\% & 2.2\% & 2.2\% & 2.2\% & 2.2\% & 2.1\% & 2.1\% & 2.0\% & 2.0\% & 1.9\% & 1.9\% & 1.8\% & \url{http://aidsinfo.unaids.org/}\\
  Georgia & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.3\% & 0.3\% & 0.3\% & 0.4\% & 0.4\% & \url{http://aidsinfo.unaids.org/}\\
  Ghana & 1.5\% & 1.7\% & 1.9\% & 2.2\% & 2.4\% & 2.6\% & 2.7\% & 2.8\% & 2.9\% & 2.9\% & 2.9\% & 2.8\% & 2.7\% & 2.6\% & 2.5\% & 2.4\% & 2.3\% & 2.2\% & 2.1\% & 2.0\% & 1.9\% & 1.9\% & 1.8\% & 1.7\% & 1.7\% & 1.6\% & \url{http://aidsinfo.unaids.org/}\\
  Greece & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.3\% & \url{http://aidsinfo.unaids.org/}\\
  Guatemala & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & 0.1\% & 0.2\% & 0.2\% & 0.2\% & 0.3\% & 0.3\% & 0.4\% & 0.4\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.6\% & 0.6\% & 0.6\% & 0.6\% & 0.6\% & 0.6\% & 0.6\% & \url{http://aidsinfo.unaids.org/}\\
  Guinea & 0.5\% & 0.6\% & 0.8\% & 1.1\% & 1.3\% & 1.5\% & 1.6\% & 1.7\% & 1.8\% & 1.8\% & 1.8\% & 1.8\% & 1.8\% & 1.7\% & 1.7\% & 1.7\% & 1.7\% & 1.6\% & 1.6\% & 1.6\% & 1.6\% & 1.6\% & 1.6\% & 1.6\% & 1.6\% & 1.6\% & \url{http://aidsinfo.unaids.org/}\\
  Guyana & --- & 0.1\% & 0.2\% & 0.3\% & 0.4\% & 0.5\% & 0.7\% & 0.8\% & 0.9\% & 0.9\% & 1\% & 1.1\% & 1.1\% & 1.2\% & 1.2\% & 1.3\% & 1.4\% & 1.4\% & 1.4\% & 1.5\% & 1.5\% & 1.5\% & 1.5\% & 1.5\% & 1.5\% & 1.5\% & \url{http://aidsinfo.unaids.org/}\\
  Haiti & 2.2\% & 2.7\% & 3.2\% & 3.7\% & 4.2\% & 4.6\% & 5.0\% & 5.2\% & 5.4\% & 5.5\% & 5.5\% & 5.4\% & 5.2\% & 5.0\% & 4.7\% & 4.4\% & 4.0\% & 3.6\% & 3.3\% & 3.0\% & 2.8\% & 2.5\% & 2.3\% & 2.0\% & 1.9\% & 1.7\% & \url{http://aidsinfo.unaids.org/}\\
  Honduras & 1.8\% & 1.9\% & 1.9\% & 2.0\% & 1.9\% & 1.9\% & 1.8\% & 1.7\% & 1.5\% & 1.4\% & 1.3\% & 1.1\% & 1\% & 0.9\% & 0.8\% & 0.8\% & 0.7\% & 0.6\% & 0.6\% & 0.5\% & 0.5\% & 0.5\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & \url{http://aidsinfo.unaids.org/}\\
  India & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & 0.1\% & 0.2\% & 0.2\% & 0.3\% & 0.3\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & --- & --- & --- & \url{http://www.unaids.org/en/resources/campaigns/globalreport2013/globalreport}\\
  Indonesia & --- & --- & --- & --- & 0.0017\% & 0.0034\% & 0.0063\% & 0.011\% & 0.017\% & 0.024\% & 0.035\% & 0.052\% & 0.079\% & 0.12\% & 0.15\% & 0.19\% & 0.23\% & 0.26\% & 0.3\% & 0.33\% & 0.37\% & 0.39\% & 0.42\% & 0.44\% & 0.46\% & 0.49\% & \url{http://aidsinfo.unaids.org/}, PLHIV 15+/ Population 15-49\\
  Iran & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & \url{http://aidsinfo.unaids.org/}\\
  Italy & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & 0.1\% & 0.2\% & 0.2\% & 0.2\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & \url{http://aidsinfo.unaids.org/}\\
  Jamaica & 0.9\% & 1.2\% & 1.5\% & 1.7\% & 1.9\% & 2.0\% & 2.1\% & 2.2\% & 2.2\% & 2.3\% & 2.2\% & 2.2\% & 2.2\% & 2.1\% & 2.0\% & 2.0\% & 1.9\% & 1.9\% & 1.8\% & 1.8\% & 1.7\% & 1.7\% & 1.7\% & 1.7\% & 1.6\% & 1.6\% & \url{http://aidsinfo.unaids.org/}\\
  Kazakhstan & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.2\% & 0.2\% & 0.2\% & \url{http://aidsinfo.unaids.org/}\\
  Kenya & 6.4\% & 8.0\% & 9.5\% & 10.9\% & 11.9\% & 12.6\% & 12.8\% & 12.7\% & 12.2\% & 11.6\% & 10.7\% & 9.9\% & 9.0\% & 8.3\% & 7.6\% & 7.0\% & 6.6\% & 6.4\% & 6.2\% & 6.1\% & 6.1\% & 6.0\% & 6.0\% & 6.0\% & 6.0\% & 5.9\% & \url{http://aidsinfo.unaids.org/}\\
  Kyrgyzstan & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & \url{http://aidsinfo.unaids.org/}\\
  Laos & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & --- & \url{http://data.worldbank.org/indicator/SH.DYN.AIDS.ZS}\\
  Lebanon & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & 0.056\% & 0.059\% & 0.061\% & 0.063\% & 0.061\% & 0.064\% & 0.063\% & 0.067\% & 0.069\% & 0.071\% & 0.074\% & 0.081\% & 0.085\% & 0.092\% & 0.094\% & \url{http://aidsinfo.unaids.org/}, PLHIV 15+/ Population 15-49\\
  Lesotho & 2.6\% & 3.6\% & 4.9\% & 6.8\% & 9.0\% & 11.5\% & 13.7\% & 15.6\% & 17.3\% & 18.7\% & 19.9\% & 20.9\% & 21.7\% & 22.2\% & 22.6\% & 22.8\% & 23.0\% & 23.1\% & 23.2\% & 23.2\% & 23.1\% & 23.1\% & 23.0\% & 22.9\% & 22.8\% & 22.7\% & \url{http://aidsinfo.unaids.org/}\\
  Liberia & 0.7\% & 0.9\% & 1.1\% & 1.4\% & 1.6\% & 1.8\% & 2.0\% & 2.2\% & 2.4\% & 2.4\% & 2.5\% & 2.5\% & 2.4\% & 2.3\% & 2.2\% & 2.1\% & 1.9\% & 1.8\% & 1.7\% & 1.5\% & 1.4\% & 1.3\% & 1.3\% & 1.2\% & 1.1\% & 1.1\% & \url{http://aidsinfo.unaids.org/}\\
  Madagascar & $<$0.1\% & $<$0.1\% & $<$0.1\% & 0.1\% & 0.2\% & 0.2\% & 0.3\% & 0.4\% & 0.5\% & 0.5\% & 0.6\% & 0.6\% & 0.6\% & 0.6\% & 0.5\% & 0.5\% & 0.5\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.4\% & \url{http://aidsinfo.unaids.org/}\\
  Malawi & 8.8\% & 10.4\% & 12.0\% & 13.3\% & 14.5\% & 15.4\% & 16.0\% & 16.5\% & 16.7\% & 16.7\% & 16.6\% & 16.3\% & 15.8\% & 15.2\% & 14.5\% & 13.9\% & 13.2\% & 12.6\% & 12.1\% & 11.6\% & 11.2\% & 10.8\% & 10.3\% & 9.9\% & 9.5\% & 9.1\% & \url{http://aidsinfo.unaids.org/}\\
  Malaysia & $<$0.1\% & 0.3\% & 0.4\% & 0.5\% & 0.6\% & 0.6\% & 0.6\% & 0.7\% & 0.7\% & 0.7\% & 0.7\% & 0.7\% & 0.7\% & 0.7\% & 0.6\% & 0.6\% & 0.6\% & 0.6\% & 0.6\% & 0.6\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.4\% & 0.4\% & \url{http://aidsinfo.unaids.org/}\\
  Mali & 1.8\% & 2.0\% & 2.1\% & 2.2\% & 2.2\% & 2.2\% & 2.1\% & 2.0\% & 1.9\% & 1.8\% & 1.7\% & 1.6\% & 1.5\% & 1.5\% & 1.4\% & 1.4\% & 1.3\% & 1.3\% & 1.3\% & 1.3\% & 1.3\% & 1.3\% & 1.3\% & 1.3\% & 1.3\% & 1.3\% & \url{http://aidsinfo.unaids.org/}\\
  Mauritania & 0.3\% & 0.3\% & 0.4\% & 0.4\% & 0.5\% & 0.6\% & 0.6\% & 0.7\% & 0.7\% & 0.8\% & 0.8\% & 0.9\% & 0.9\% & 0.9\% & 0.9\% & 0.9\% & 0.9\% & 0.8\% & 0.8\% & 0.8\% & 0.7\% & 0.7\% & 0.7\% & 0.6\% & 0.6\% & 0.6\% & \url{http://aidsinfo.unaids.org/}\\
  Mauritius & --- & --- & --- & 0.1\% & 0.2\% & 0.3\% & 0.4\% & 0.6\% & 0.7\% & 0.9\% & 1\% & 1.1\% & 1.1\% & 1.2\% & 1.2\% & 1.2\% & 1.2\% & 1.2\% & 1.2\% & 1.1\% & 1.1\% & 1\% & 1\% & 1\% & 0.9\% & 0.9\% & \url{http://aidsinfo.unaids.org/}\\
  Mexico & 0.1\% & 0.2\% & 0.2\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & \url{http://aidsinfo.unaids.org/}\\
  Mongolia & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & 0.019\% & 0.025\% & 0.033\% & 0.04\% & 0.054\% & 0.067\% & 0.081\% & 0.097\% & 0.11\% & --- & \url{https://www.cia.gov/library/publications/the-world-factbook/rankorder/2155rank.html}, Reported \# HIV cases (total) / Population (15-49)\\
  Morocco & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & \url{http://aidsinfo.unaids.org/}\\
  Mozambique & 1.6\% & 2.0\% & 2.4\% & 2.9\% & 3.5\% & 4.1\% & 4.8\% & 5.5\% & 6.3\% & 7.1\% & 7.8\% & 8.6\% & 9.2\% & 9.8\% & 10.2\% & 10.5\% & 10.8\% & 10.9\% & 11.1\% & 11.2\% & 11.2\% & 11.1\% & 11.0\% & 10.9\% & 10.7\% & 10.5\% & \url{http://aidsinfo.unaids.org/}\\
  Myanmar & --- & 0.1\% & 0.2\% & 0.3\% & 0.3\% & 0.4\% & 0.4\% & 0.5\% & 0.6\% & 0.7\% & 0.8\% & 0.8\% & 0.9\% & 0.9\% & 0.9\% & 0.9\% & 0.9\% & 0.9\% & 0.9\% & 0.9\% & 0.8\% & 0.8\% & 0.8\% & 0.8\% & 0.8\% & 0.8\% & \url{http://aidsinfo.unaids.org/}\\
  Namibia & 1.4\% & 2.2\% & 3.3\% & 4.8\% & 6.8\% & 9.1\% & 11.4\% & 13.5\% & 15.1\% & 16.3\% & 17.1\% & 17.4\% & 17.4\% & 17.2\% & 16.8\% & 16.3\% & 15.9\% & 15.5\% & 15.3\% & 15.0\% & 14.7\% & 14.4\% & 14.1\% & 13.8\% & 13.6\% & 13.3\% & \url{http://aidsinfo.unaids.org/}\\
  Nepal & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.2\% & 0.2\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.2\% & 0.2\% & 0.2\% & \url{http://aidsinfo.unaids.org/}\\
  Netherlands & 0.095\% & 0.12\% & 0.14\% & 0.17\% & 0.19\% & 0.19\% & 0.2\% & 0.21\% & 0.22\% & 0.22\% & 0.24\% & 0.24\% & 0.25\% & 0.26\% & 0.27\% & 0.28\% & 0.29\% & 0.31\% & 0.31\% & 0.31\% & 0.33\% & 0.34\% & 0.35\% & --- & --- & --- & \url{http://www.unaids.org/en/resources/campaigns/globalreport2013/globalreport}\\
  New Zealand & 0.06\% & 0.06\% & 0.06\% & 0.06\% & 0.06\% & 0.06\% & 0.06\% & 0.06\% & 0.06\% & 0.05\% & 0.05\% & 0.05\% & 0.05\% & 0.05\% & 0.05\% & 0.05\% & 0.05\% & 0.05\% & --- & --- & --- & --- & --- & --- & 0.091\% & 0.15\% & \url{http://dnmeds.otago.ac.nz/departments/psm/research/aids/pdf/75%20AIDS-NZ%20May%202016.pdf}, Estimated PLHIV (total) / Population (15-49)\\
  Nicaragua & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & \url{http://aidsinfo.unaids.org/}\\
  Niger & 0.2\% & 0.3\% & 0.4\% & 0.5\% & 0.6\% & 0.7\% & 0.9\% & 1\% & 1.1\% & 1.2\% & 1.2\% & 1.2\% & 1.2\% & 1.2\% & 1.2\% & 1.1\% & 1\% & 0.9\% & 0.8\% & 0.8\% & 0.7\% & 0.6\% & 0.6\% & 0.5\% & 0.5\% & 0.5\% & \url{http://aidsinfo.unaids.org/}\\
  Nigeria & 1.4\% & 1.6\% & 1.9\% & 2.2\% & 2.5\% & 2.9\% & 3.2\% & 3.5\% & 3.7\% & 3.8\% & 3.9\% & 4.0\% & 4.0\% & 4.0\% & 3.9\% & 3.8\% & 3.8\% & 3.7\% & 3.6\% & 3.5\% & 3.5\% & 3.4\% & 3.3\% & 3.2\% & 3.2\% & 3.1\% & \url{http://aidsinfo.unaids.org/}\\
  Pakistan & 0.0047\% & 0.0054\% & 0.0063\% & 0.007\% & 0.0076\% & 0.0083\% & 0.0089\% & 0.0098\% & 0.011\% & 0.012\% & 0.013\% & 0.015\% & 0.016\% & 0.018\% & 0.022\% & 0.025\% & 0.028\% & 0.032\% & 0.037\% & 0.046\% & 0.054\% & 0.063\% & 0.072\% & 0.082\% & 0.092\% & 0.1\% & \url{http://aidsinfo.unaids.org/}, PLHIV 15+/ Population 15-49\\
  Panama & 0.3\% & 0.4\% & 0.5\% & 0.6\% & 0.7\% & 0.7\% & 0.7\% & 0.7\% & 0.7\% & 0.7\% & 0.7\% & 0.7\% & 0.7\% & 0.6\% & 0.6\% & 0.6\% & 0.6\% & 0.6\% & 0.7\% & 0.7\% & 0.7\% & 0.7\% & 0.7\% & 0.7\% & 0.7\% & 0.7\% & \url{http://aidsinfo.unaids.org/}\\
  Papua New Guinea & $<$0.1\% & $<$0.1\% & 0.1\% & 0.2\% & 0.3\% & 0.3\% & 0.4\% & 0.5\% & 0.6\% & 0.7\% & 0.8\% & 0.9\% & 0.9\% & 0.9\% & 0.9\% & 0.9\% & 0.9\% & 0.9\% & 0.9\% & 0.8\% & 0.8\% & 0.8\% & 0.8\% & 0.8\% & 0.8\% & 0.8\% & \url{http://aidsinfo.unaids.org/}\\
  Paraguay & 0.1\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & \url{http://aidsinfo.unaids.org/}\\
  Peru & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & \url{http://aidsinfo.unaids.org/}\\
  Philippines & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & 0.006\% & 0.0071\% & 0.0085\% & 0.01\% & 0.012\% & 0.014\% & 0.017\% & 0.02\% & 0.024\% & 0.03\% & 0.039\% & 0.047\% & 0.063\% & 0.071\% & --- & \url{http://www.unaids.org/sites/default/files/country/documents//PHL_narrative_report_2014.pdf}, Projected PLHIV / Population (15-49)\\
  Moldova & --- & --- & --- & --- & --- & --- & 0.1\% & 0.1\% & 0.2\% & 0.2\% & 0.3\% & 0.3\% & 0.3\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.5\% & 0.5\% & 0.5\% & 0.6\% & 0.6\% & 0.6\% & 0.6\% & 0.6\% & \url{http://aidsinfo.unaids.org/}\\
  Russia & --- & --- & --- & --- & --- & --- & --- & 0.0069\% & 0.012\% & 0.038\% & 0.12\% & 0.23\% & 0.3\% & 0.35\% & 0.4\% & 0.45\% & 0.51\% & 0.57\% & 0.63\% & --- & --- & *1.385 & --- & --- & --- & 2.2\% & \url{http://data.unaids.org/pub/Report/2008/russia_2008_country_progress_report_en.pdf}, \url{http://www.hivrussia.ru/skm/info_en.shtml}, \url{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4247269/table/T1/}, \url{http://www.unaids.org/sites/default/files/country/documents/RUS_narrative_report_2015.pdf}\\
  Rwanda & 5.2\% & 5.7\% & 6.1\% & 6.3\% & 6.3\% & 6.3\% & 6.1\% & 5.9\% & 5.6\% & 5.3\% & 4.9\% & 4.6\% & 4.2\% & 3.9\% & 3.6\% & 3.4\% & 3.3\% & 3.3\% & 3.2\% & 3.2\% & 3.1\% & 3.1\% & 3.1\% & 3.0\% & 3.0\% & 2.9\% & \url{http://aidsinfo.unaids.org/}\\
  Saint Lucia & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & 0.28\% & --- & --- & 0.56\% & 0.46\% & --- & 0.69\% & --- & \url{http://www.unaids.org/sites/default/files/country/documents/file,68394,es..pdf}, \url{http://www.unaids.org/sites/default/files/country/documents/LCA_narrative_report_2015.pdf}\\
  Senegal & 0.2\% & 0.2\% & 0.2\% & 0.3\% & 0.3\% & 0.4\% & 0.4\% & 0.5\% & 0.6\% & 0.7\% & 0.7\% & 0.8\% & 0.9\% & 0.9\% & 0.9\% & 0.9\% & 0.8\% & 0.8\% & 0.8\% & 0.7\% & 0.7\% & 0.7\% & 0.6\% & 0.6\% & 0.6\% & 0.5\% & \url{http://aidsinfo.unaids.org/}\\
  Sierra Leone & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.2\% & 0.3\% & 0.4\% & 0.6\% & 0.7\% & 0.9\% & 1.1\% & 1.3\% & 1.4\% & 1.5\% & 1.6\% & 1.7\% & 1.7\% & 1.7\% & 1.7\% & 1.6\% & 1.6\% & 1.5\% & 1.5\% & 1.4\% & 1.3\% & \url{http://aidsinfo.unaids.org/}\\
  Somalia & $<$0.1\% & 0.1\% & 0.2\% & 0.2\% & 0.3\% & 0.4\% & 0.4\% & 0.5\% & 0.6\% & 0.6\% & 0.6\% & 0.6\% & 0.6\% & 0.6\% & 0.6\% & 0.6\% & 0.6\% & 0.6\% & 0.6\% & 0.6\% & 0.6\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & \url{http://aidsinfo.unaids.org/}\\
  South Africa & 1\% & 1.6\% & 2.6\% & 3.9\% & 5.7\% & 7.9\% & 10.3\% & 12.8\% & 15.1\% & 17.1\% & 18.7\% & 19.9\% & 20.6\% & 21.0\% & 21.0\% & 20.8\% & 20.4\% & 19.9\% & 19.5\% & 19.2\% & 18.9\% & 18.8\% & 18.8\% & 18.9\% & 19.0\% & 19.2\% & \url{http://aidsinfo.unaids.org/}\\
  South Sudan & 0.1\% & 0.3\% & 0.4\% & 0.6\% & 0.9\% & 1.1\% & 1.4\% & 1.6\% & 1.9\% & 2.2\% & 2.4\% & 2.6\% & 2.8\% & 2.9\% & 2.9\% & 2.9\% & 2.9\% & 2.8\% & 2.8\% & 2.8\% & 2.7\% & 2.7\% & 2.6\% & 2.6\% & 2.5\% & 2.5\% & \url{http://aidsinfo.unaids.org/}\\
  Spain & 0.4\% & 0.4\% & 0.4\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & \url{http://aidsinfo.unaids.org/}\\
  Sri Lanka & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & 0.038\% & --- & 0.028\% & --- & 0.023\% & 0.027\% & 0.029\% & 0.032\% & --- & \url{http://www.unaids.org/sites/default/files/country/documents/LKA_narrative_report_2016.pdf}, Estimated \# PLHIV (total) / Population (15-49)\\
  Sudan & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.3\% & 0.3\% & \url{http://aidsinfo.unaids.org/}\\
  Suriname & 0.1\% & 0.2\% & 0.3\% & 0.3\% & 0.4\% & 0.5\% & 0.6\% & 0.8\% & 0.9\% & 1\% & 1.1\% & 1.2\% & 1.3\% & 1.3\% & 1.3\% & 1.3\% & 1.3\% & 1.2\% & 1.2\% & 1.2\% & 1.2\% & 1.2\% & 1.1\% & 1.1\% & 1.1\% & 1.1\% & \url{http://aidsinfo.unaids.org/}\\
  Swaziland & 2.2\% & 4.0\% & 6.5\% & 9.4\% & 12.6\% & 15.7\% & 18.5\% & 20.9\% & 22.7\% & 24.0\% & 24.8\% & 25.4\% & 25.6\% & 25.7\% & 25.7\% & 25.7\% & 25.8\% & 26.1\% & 26.4\% & 26.7\% & 27.2\% & 27.6\% & 28.0\% & 28.4\% & 28.7\% & 28.8\% & \url{http://aidsinfo.unaids.org/}\\
  Sweden & 0.089\% & 0.11\% & 0.13\% & 0.14\% & 0.15\% & 0.16\% & 0.17\% & 0.17\% & 0.18\% & 0.18\% & 0.19\% & 0.19\% & 0.19\% & 0.21\% & 0.21\% & 0.21\% & 0.22\% & 0.22\% & 0.22\% & 0.23\% & 0.23\% & 0.23\% & 0.23\% & --- & --- & --- & \url{http://www.unaids.org/en/resources/campaigns/globalreport2013/globalreport}\\
  Switzerland & 0.26\% & 0.29\% & 0.31\% & 0.33\% & 0.34\% & 0.34\% & 0.33\% & 0.35\% & 0.36\% & 0.37\% & 0.38\% & 0.39\% & 0.4\% & 0.43\% & 0.44\% & 0.46\% & 0.46\% & 0.48\% & 0.49\% & 0.51\% & 0.52\% & 0.53\% & 0.56\% & --- & --- & --- & \url{http://www.unaids.org/en/resources/campaigns/globalreport2013/globalreport}\\
  Tajikistan & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & 0.1\% & 0.1\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & \url{http://aidsinfo.unaids.org/}\\
  Thailand & 0.5\% & 0.9\% & 1.3\% & 1.6\% & 1.8\% & 1.9\% & 2.0\% & 1.9\% & 1.9\% & 1.8\% & 1.7\% & 1.7\% & 1.6\% & 1.5\% & 1.4\% & 1.4\% & 1.4\% & 1.3\% & 1.3\% & 1.3\% & 1.3\% & 1.3\% & 1.2\% & 1.2\% & 1.2\% & 1.1\% & \url{http://aidsinfo.unaids.org/}\\
  Timor-Leste & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & 0.013\% & --- & --- & --- & --- & --- & --- & --- & --- & --- & 0.085\% & 0.097\% & \url{http://www.hivpolicy.org/Library/HPP000476.pdf}, Estimated \# PLHIV (total) / Population (15-49)\\
  Togo & 1.2\% & 1.5\% & 1.9\% & 2.3\% & 2.8\% & 3.2\% & 3.6\% & 4.0\% & 4.4\% & 4.6\% & 4.8\% & 4.9\% & 4.9\% & 4.8\% & 4.6\% & 4.4\% & 4.1\% & 3.9\% & 3.6\% & 3.3\% & 3.1\% & 2.9\% & 2.8\% & 2.6\% & 2.5\% & 2.4\% & \url{http://aidsinfo.unaids.org/}\\
  Trinidad and Tobago & 0.4\% & 0.5\% & 0.5\% & 0.6\% & 0.7\% & 0.8\% & 0.8\% & 0.9\% & 0.9\% & 1\% & 1\% & 1\% & 1.1\% & 1.1\% & 1.1\% & 1.1\% & 1.2\% & 1.2\% & 1.2\% & 1.2\% & 1.2\% & 1.2\% & 1.2\% & 1.2\% & 1.2\% & 1.2\% & \url{http://aidsinfo.unaids.org/}\\
  Tunisia & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & 0.019\% & 0.022\% & 0.024\% & 0.027\% & 0.03\% & 0.033\% & 0.037\% & 0.04\% & 0.043\% & \url{http://aidsinfo.unaids.org/}, PLHIV 15+/ Population 15-49\\
  Uganda & 13.1\% & 13.3\% & 13.1\% & 12.7\% & 12.1\% & 11.4\% & 10.6\% & 9.8\% & 9.0\% & 8.4\% & 7.8\% & 7.3\% & 7.0\% & 6.7\% & 6.5\% & 6.5\% & 6.6\% & 6.6\% & 6.7\% & 6.9\% & 7.0\% & 7.1\% & 7.2\% & 7.2\% & 7.1\% & 7.1\% & \url{http://aidsinfo.unaids.org/}\\
  Ukraine & --- & --- & --- & 0.1\% & 0.1\% & 0.2\% & 0.3\% & 0.3\% & 0.4\% & 0.6\% & 0.7\% & 0.8\% & 0.8\% & 0.9\% & 0.9\% & 1\% & 1\% & 0.9\% & 0.9\% & 0.9\% & 0.9\% & 0.8\% & 0.8\% & 0.8\% & 0.8\% & 0.9\% & \url{http://aidsinfo.unaids.org/}\\
  United Kingdom & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & 0.096\% & 0.11\% & 0.12\% & 0.13\% & 0.15\% & 0.16\% & 0.17\% & 0.17\% & 0.18\% & 0.19\% & 0.19\% & 0.2\% & --- & --- & \url{https://www.gov.uk/government/uploads/system/uploads/attachment_data/file/335452/HIV_annual_report_2012.pdf}, Estimated \# PLHIV (15-49) / Population (15-49)\\
  Tanzania & 5.1\% & 5.8\% & 6.4\% & 7.0\% & 7.4\% & 7.7\% & 7.9\% & 8.0\% & 7.9\% & 7.8\% & 7.7\% & 7.5\% & 7.3\% & 7.1\% & 6.8\% & 6.5\% & 6.3\% & 6.1\% & 6.0\% & 5.8\% & 5.6\% & 5.4\% & 5.2\% & 5.1\% & 4.9\% & 4.7\% & \url{http://aidsinfo.unaids.org/}\\
  United States & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & --- & 0.21\% & 0.22\% & 0.23\% & 0.24\% & 0.25\% & 0.26\% & 0.26\% & 0.37\% & 0.37\% & 0.37\% & 0.37\% & 0.37\% & --- & --- & --- & \url{http://www.cdc.gov/hiv/library/reports/surveillance/pastissues.html}, Estimated \# PLHIV (15-49) / Population (15-49)\\
  Uruguay & 0.1\% & 0.1\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.4\% & 0.4\% & 0.4\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & \url{http://aidsinfo.unaids.org/}\\
  Uzbekistan & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & 0.1\% & 0.2\% & 0.2\% & 0.3\% & 0.3\% & 0.4\% & 0.4\% & 0.4\% & 0.3\% & 0.3\% & 0.3\% & 0.3\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & 0.2\% & \url{http://aidsinfo.unaids.org/}\\
  Venezuela & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & $<$0.1\% & 0.1\% & 0.2\% & 0.2\% & 0.3\% & 0.3\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & \url{http://aidsinfo.unaids.org/}\\
  Viet Nam & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.1\% & 0.2\% & 0.2\% & 0.3\% & 0.3\% & 0.3\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.4\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & 0.5\% & \url{http://aidsinfo.unaids.org/}\\
  Yemen & --- & 0.022\% & 0.024\% & 0.026\% & 0.028\% & 0.032\% & 0.032\% & 0.034\% & 0.034\% & 0.036\% & 0.037\% & 0.039\% & 0.04\% & 0.042\% & 0.043\% & 0.044\% & 0.047\% & 0.048\% & 0.05\% & 0.052\% & 0.054\% & 0.056\% & 0.058\% & 0.06\% & 0.063\% & 0.065\% & \url{http://aidsinfo.unaids.org/}, PLHIV 15+/ Population 15-49\\
  Zambia & 12.4\% & 13.5\% & 14.3\% & 14.9\% & 15.3\% & 15.6\% & 15.7\% & 15.8\% & 15.7\% & 15.6\% & 15.4\% & 15.2\% & 14.9\% & 14.6\% & 14.4\% & 14.1\% & 13.8\% & 13.6\% & 13.6\% & 13.5\% & 13.5\% & 13.4\% & 13.3\% & 13.2\% & 13.1\% & 12.9\% & \url{http://aidsinfo.unaids.org/}\\
  Zimbabwe & 15.0\% & 17.7\% & 20.3\% & 22.5\% & 24.2\% & 25.3\% & 26.0\% & 26.0\% & 25.7\% & 24.9\% & 24.0\% & 22.8\% & 21.6\% & 20.4\% & 19.2\% & 18.1\% & 17.2\% & 16.5\% & 16.0\% & 15.7\% & 15.5\% & 15.3\% & 15.2\% & 15.0\% & 14.9\% & 14.7\% & \url{http://aidsinfo.unaids.org/}\\
  \hline
  \multicolumn{1}{l}{NOTES}
\end{longtable}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "supplementary_text"
%%% End: