### 2016-10-19 Documents/extended_data_fig_2.tex
More changes for Nature.

\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage{microtype}
\usepackage{geometry}

\usepackage{graphicx}

\usepackage[font={sf}, labelfont={bf},
            labelsep=space, singlelinecheck=off]{caption}
\usepackage[font={sf}]{floatrow}

\usepackage[pdfborder={0 0 0}]{hyperref}

\renewcommand{\figurename}{Extended Data Fig.}
\setcounter{figure}{1}

\begin{document}

\begin{figure}
  \centering
  \includegraphics{../Codes/plots/prcc.pdf}
  \caption{Sensitivity of global infections averted from 2015 to 2035
    to model parameters.  Left is the sensitivity of infections
    averted by 95--95--95, relative to status quo.  Right is the
    sensitivity of vaccination with status quo levels of diagnosis and
    treatment, relative to status quo.  Sensitivity is measured by the
    partial rank correlation coefficient (PRCC).}
  \label{PRCCs}
\end{figure}

\end{document}