### 2016-11-30 Documents/extended_data_fig_2.tex
Cropped paper sizes for extended data figs.

\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage{microtype}
\usepackage[scale = 0.95,
            paperwidth = 6.3in,
            paperheight = 7.7in]{geometry}

\usepackage{graphicx}

\usepackage[font={sf}, labelfont={bf},
            labelsep=space, singlelinecheck=off]{caption}
\usepackage[font={sf}]{floatrow}

\usepackage[pdfborder={0 0 0}]{hyperref}

\renewcommand{\figurename}{Extended Data Fig.}
\setcounter{figure}{1}

\pagestyle{empty}

\begin{document}

\begin{figure}
   \centering
   \includegraphics{../Codes/plots/prcc.pdf}
   \caption{Sensitivity of global HIV incidence from 2015 to 2035 to
     model parameters.  Left is the sensitivity of infections averted
     by 95--95--95, relative to status quo.  Right is the sensitivity
     of vaccination with status quo levels of diagnosis and treatment,
     relative to status quo.  Sensitivity is measured by the partial
     rank correlation coefficient (PRCC).}
   \label{PRCCs}
\end{figure}

\end{document}