### 2016-12-15 docs/figs.tex
More tweaks for PNAS.

\documentclass{article}

\usepackage{common}

\usepackage[textwidth=505.5bp]{geometry}

\usepackage{graphics}

\renewcommand{\figurename}{Fig.}

\begin{document}

\begin{figure}
  \centering
  \includegraphics{../src/plots/infections_averted_map.pdf}
  \caption{Infections averted between 2015 and 2035, compared to
    maintaining status quo diagnosis, treatment, and viral
    suppression, by the UNAIDS 95–95–95 target (top); rollout of
    vaccination in 2020 while maintaining status quo levels of
    diagnosis, treatment, and viral suppression (middle); and the
    UNAIDS 95–95–95 target combined with 2020 vaccination rollout
    (bottom). Predictions are medians over 1000 model simulations in
    which parameter values were sampled from data-driven
    distributions. See Figs. S4–S140 for projections and uncertainty
    estimation globally, by region, and by country.}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{../src/plots/effectiveness.pdf}
  \caption{Effectiveness of HIV interventions globally (left column)
    and for selected countries in terms of cumulative incidence (top
    row), annual incidence per million population (2nd row), PLHIV
    (3rd row), and cumulative HIV mortality (bottom row). Predictions
    are medians over 1000 model simulations in which parameter values
    were sampled from data-driven distributions. See Figs. S4–S140 for
    median and uncertainty estimates of effectiveness of HIV
    interventions for all 127 countries.}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{../src/plots/effectiveness_regions.pdf}
  \caption{Effectiveness of HIV interventions for UNAIDS regions in
    terms of cumulative incidence from 2015 to 2035 (top) and number
    of PLHIV in 2035 (bottom). Medians (colored bars) and 1st and 3rd
    quartiles (black error bars) were generated from 1000 model
    simulations in which parameter values were sampled from
    data-driven distributions. See Figs. S4–S140 for median and
    uncertainty estimates of the effectiveness of HIV interventions
    for UNAIDS regions.}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{../src/plots/vaccine_scenarios.pdf}
  \caption{Scenario analysis for vaccination under six strategies,
    globally (left column) and for selected countries (remaining
    columns), with variation in vaccine efficacy, ultimate coverage,
    initiation date of vaccination rollout, and rate of
    scale-up. Vaccine scenario projections are generated from
    simulations parametrized with modal values.}
\end{figure}

\end{document}